---
title: "MalariaVax Data Processing"
author: "Christine Davis"
date: "1/4/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

The systemsseRology package is dependent on ropls and must be downloaded from 
github. Please run the following in your console prior to utilizing this Rmd.

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ropls")
install.packages("devtools")
library(devtools)
install_github("LoosC/systemsseRology", ref = "reboot")


Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

Load packages
```{r load}
library(systemsseRology)
library(readxl)
library(hablar)
library(pheatmap)
library(DMwR)
library(RColorBrewer)
library(summarytools)
library(ggpubr)
library(tidyr)
library(ggplot2)
library(ggforce)
library(ggsignif)
library(tidyverse)
library(ropls)
library(corrr)
library(igraph)
library(ggraph)
library(ggrepel)
set.seed(0)  # random seed for reproducibility
```

# Classification and preprocessing

Read the data from an Excel file. Z-score the data and assign the label in the variable y.

Data includes 37 parameters and 44 samples, divided into 4 classes (vaccine regimes 1-4, with regime 3 being DFx)
Group labels (y, in PCA): column 3
X-data: columns 7-70
Data pre-processing performed to: 
1) exclude measurements (blue starred measurements in data spreadsheet) --> 
processed as NA, warnings suppressed
2) average replicates (23 assays have duplicates, 1 assay has 4-peat)

```{r data}
#Load data
file_name <- "../MalariaVax/RH5 VAC063 G1-4 systems serology.xlsx"
data <- as.data.frame(read_excel(file_name, sheet = "UPDATED Subset Data"))
X <- suppressWarnings(
  as.matrix(as.data.frame(lapply(data[, 7:ncol(data)], as.numeric))))
data_new <- data[,7:19] #Pull glycan data, which has no replicates to average

#Average all duplicate measurements and rename
data_new$"IgG" <- rowMeans(X[,c(14,15)],na.rm = TRUE)
data_new$"IgM" <- rowMeans(X[,c(16,17)],na.rm = TRUE)
data_new$"IgA1" <- rowMeans(X[,c(18,19)],na.rm = TRUE)
data_new$"IgA2" <- rowMeans(X[,c(20,21)],na.rm = TRUE)
data_new$"IgG1" <- rowMeans(X[,c(22,23)],na.rm = TRUE)
data_new$"IgG3" <- rowMeans(X[,c(24,25)],na.rm = TRUE)
data_new$"IgG2" <- rowMeans(X[,c(26,27)],na.rm = TRUE)
data_new$"IgG4" <- rowMeans(X[,c(28,29)],na.rm = TRUE)
data_new$"FcR2AH" <- rowMeans(X[,c(30,31)],na.rm = TRUE)
data_new$"FcR2AR" <- rowMeans(X[,c(32,33)],na.rm = TRUE)
data_new$"FcR3AF" <- rowMeans(X[,c(34,35)],na.rm = TRUE)
data_new$"FcR3AV" <- rowMeans(X[,c(36,37)],na.rm = TRUE)
data_new$"FcR2B" <- rowMeans(X[,c(38,39)],na.rm = TRUE)
data_new$"FcR3B" <- rowMeans(X[,c(40,41)],na.rm = TRUE)
data_new$"FcaR" <- rowMeans(X[,c(42,43)],na.rm = TRUE)
data_new$"FcRN" <- rowMeans(X[,c(44,45)],na.rm = TRUE)
data_new$"C1q" <- rowMeans(X[,c(46,47)],na.rm = TRUE)
data_new$"AvidityIndex" <- rowMeans(X[,c(48,49)],na.rm = TRUE)
data_new$"ADCP" <- rowMeans(X[,c(50,51)],na.rm = TRUE)
data_new$"ADCD" <- rowMeans(X[,c(52,53)],na.rm = TRUE)
data_new$"ADNP" <- rowMeans(X[,c(54,55,56,57)],na.rm = TRUE) #4 replicates ADNP
data_new$"ADNKAMIP1BwithinNKs" <- rowMeans(X[,c(58,59)],na.rm = TRUE)
data_new$"ADNKACD107awithinNKs" <- rowMeans(X[,c(60,61)],na.rm = TRUE)
data_new$"ADNKAIFNgwithinNKs" <- rowMeans(X[,c(62,63)],na.rm = TRUE)

#Ensure all data is numeric form
X <- suppressWarnings(as.matrix(as.data.frame(lapply(data_new, as.numeric))))

#Rename columns to be more usable
evens <- seq(2,26,2)
colnames(X)[1:13] <- unlist(strsplit(colnames(X)[1:13],"...",fixed = T))[evens]
colnames(X) <- gsub(".","", fixed = TRUE, colnames(X))

## Remove columns with more than 50% NA, then use k-nearest neighbors on 
# ONLY multivariate analysis data to impute missing values
rownames(X) <- data[,1] #trial subject numbers
Ximp <- X[, which(colMeans(!is.na(X)) > 0.5)]
Ximp <- impute.knn(Ximp, k = 3)#knnImputation(Ximp, k=3)
Ximp <- Ximp$data
#no imputation for data intended for univariate analysis
X_uni <-X[, which(colMeans(!is.na(X)) > 0.5)] 
```

#Optional normalization section
Only run these if you plan to normalize to individual titers or group means
Skip to "Finish preprocessing, create heatmap" if not.

Check if distribution is normal if you plan to normalize by mean or 
median of group. Can use mean if normally distributed, but otherwise should
consider median.

```{r distribution check}
#Visually examine distributions
ggdensity(Ximp[which(grepl("1", (data[, 3]))),11], 
          main = "Density plot of IgG Group 1",
          xlab = "IgG")
ggdensity(Ximp[which(grepl("2", (data[, 3]))),11], 
          main = "Density plot of IgG Group 2",
          xlab = "IgG")
ggdensity(Ximp[which(grepl("3", (data[, 3]))),11], 
          main = "Density plot of IgG Group 3",
          xlab = "IgG")
ggdensity(Ximp[which(grepl("4", (data[, 3]))),11], 
          main = "Density plot of IgG Group 4",
          xlab = "IgG")
#Normality check via Shapiro-Wilk test
shapiro.test(Ximp[which(grepl("1", (data[, 3]))),11])
shapiro.test(Ximp[which(grepl("2", (data[, 3]))),11])
shapiro.test(Ximp[which(grepl("3", (data[, 3]))),11])
shapiro.test(Ximp[which(grepl("4", (data[, 3]))),11])
```

Option 1: Normalize all to IgG per individual

```{r Normalize on per-individual titer}
#Multivariate data processing
IgG <- Ximp[,11] #pull titer values
Ximp_norm <- as.matrix(Ximp)/as.numeric(t(IgG))
Ximp_norm[,11] <- IgG #replace vector of 1s with original IgG values
Ximp_norm[,1:10] <- Ximp[,1:10] #don't normalize % glycosylation by IgG

#Repeat for univariate data
IgG_uni <- X_uni[,11] 
X_uni_norm <- as.matrix(X_uni)/as.numeric(t(IgG_uni))
X_uni_norm[,11] <- IgG_uni
X_uni_norm[,1:10] <- X_uni[,1:10]
```

Option 2: this section normalizes all vaccinees within a group (1/2/3/4) 
to the mean or median of their group's IgG measurement. Change mean <-> median
based on any normality assumptions

```{r Normalize by Median/Mean IgG}

#Get group-specific rows for all features
#Multivariate data
Ximp_1 <- as.matrix(Ximp[which(grepl("1", (data[, 3]))),])
Ximp_2 <- as.matrix(Ximp[which(grepl("2", (data[, 3]))),])
Ximp_3 <- as.matrix(Ximp[which(grepl("3", (data[, 3]))),])
Ximp_4 <- as.matrix(Ximp[which(grepl("4", (data[, 3]))),])
#Univariate data
Xuni_1 <- as.matrix(X_uni[which(grepl("1", (data[, 3]))),])
Xuni_2 <- as.matrix(X_uni[which(grepl("2", (data[, 3]))),])
Xuni_3 <- as.matrix(X_uni[which(grepl("3", (data[, 3]))),])
Xuni_4 <- as.matrix(X_uni[which(grepl("4", (data[, 3]))),])

#Get mean or median of group IgG and divide all group features by it. 
#Put normalized values into Ximp_norm matrix. Swap out mean or median based on 
#normality of data calculated above with Shapiro-Wilk normality test

#Multivariate data
Ximp_1b <- Ximp_1/as.numeric(median(Ximp_1[,11]))
Ximp_2b <- Ximp_2/as.numeric(median(Ximp_2[,11]))
Ximp_3b <- Ximp_3/as.numeric(median(Ximp_3[,11]))
Ximp_4b <- Ximp_4/as.numeric(median(Ximp_4[,11]))
Ximp_norm <- rbind(Ximp_1b, Ximp_2b, Ximp_3b, Ximp_4b)

#Repeat for univariate data
Xuni_1b <- Xuni_1/as.numeric(median(Xuni_1[,11]))
Xuni_2b <- Xuni_2/as.numeric(median(Xuni_2[,11]))
Xuni_3b <- Xuni_3/as.numeric(median(Xuni_3[,11]))
Xuni_4b <- Xuni_4/as.numeric(median(Xuni_4[,11]))
X_uni_norm <- rbind(Xuni_1b, Xuni_2b, Xuni_3b, Xuni_4b)

#Replace features 1-11 with non-normalized values for glycans and IgG
X_uni_norm[,1:11] <- X_uni[,1:11]

```

Finish pre-processing, create heatmap

```{r finalize data, create heatmap}
#Comment out this section if not normalizing to titer
###
#Ximp <- scale(Ximp_norm, center = TRUE, scale = TRUE)
#X_uni <- scale(X_uni_norm, center = TRUE, scale = TRUE)
###

#Comment out this section if normalizing to titer
###
Ximp <- scale(Ximp, center = TRUE, scale = TRUE)
X_uni <- scale(X_uni, center = TRUE, scale = TRUE)
###

y <- factor(data[, 3]) #group number for each subject
names(y) <- rownames(X)

#Create heatmap annotation for coloring by group
annotation_row = data.frame(group = y)
rownames(annotation_row) <- rownames(Ximp)
my_colors <- list(
  group = c("1" = '#8C02CC', "2" = '#5192C3', "3" = '#45C902', "4" = '#045CFF')
  )

#Create heatmap of the z-scored data indicating vaccination group
p <- pheatmap(Ximp, annotation_colors = my_colors, 
              annotation_row = annotation_row, 
              cluster_cols = F, cluster_rows = F,
              cellwidth = 4, cellheight = 4, fontsize = 3, 
              border_color = "NA", treeheight_row = 3, 
              treeheight_col = 3)

#print to pdf- change to be appropriate to your own directory
pdf("../MalariaVax/Figures/heatmap.pdf", 
       width = 8, height = 5) 
print(p)
dev.off()
```

The variable df_features contains information about the features, mainly used for visualization routines. 

```{r annotation}
#Add feature names to univariate dataframe
df_features_uni <- data.frame(name = colnames(X))
df_features_uni$attribute <- factor(df_features_uni$name)

#Add feature names to multivariate dataframe
df_features <- data.frame(name = colnames(Ximp))
df_features$label <- factor(df_features$name)
df_features$feature_class <- rep(NA, length = nrow(df_features))

#Name features based on feature type
df_features$feature_class[which(grepl("G0|G1|G2", df_features$name))] <- "glycan"
df_features$feature_class[which(grepl("IgG|IgA|IgM", df_features$name))] <- "titer"
df_features$feature_class[which(grepl("FcR", df_features$name))] <- "FcR"
df_features$feature_class[which(grepl("FcaR", df_features$name))] <- "FcR"
df_features$feature_class[which(grepl("AD", df_features$name))] <- "function"
df_features$feature_class[which(grepl("Avid", df_features$name))] <- "function"
df_features$feature_class[which(grepl("C1q", df_features$name))] <- "FcR"
print(df_features)

```

#Univariate plotting

Assess feature differences between vaccinated groups 1-4 using Kruskal-Wallis 
tests, corrected for multiple testing using the Benjamini-Hochberg procedure. 
If significant differences are found, proceed to test pair-wise by Wilcoxon 
testing. Stars indicate significance as (*p < 0.05, **p < 0.01, ***p < 0.001).

Plot all features grouped by class.

```{r univariate statistics}
# Reformat the data as a long dataframe for the plotting procedure
dfBox <- reshape2::melt(cbind(as.data.frame(X_uni, group = names(y)), 
                              id.var = y))
colnames(dfBox) <- c("group", "features", "value")
#Which groups do you want to compare?
my_comparisons <- list( c("1", "2"), c("1", "3"), c("1", "4") , 
                        c("2", "3"), c("2", "4"), c("3", "4") )

#Calculate unmodified p-values using Kruskal-Wallis test
feature_list <- as.character(unique(dfBox$features))
pvalue <- rep(NA, length=length(feature_list))

#Loops over all unique features, testing via KW and pulling out/storing p-value
for (i in 1:length(pvalue))
  {
  featuredfeature <- feature_list[i]
  subset <- dfBox[dfBox$features == featuredfeature,]
  kwvalue <- subset[,3] #feature value
  kwclass <- subset[,1] #group assignment
  dfB <- kruskal.test(x = kwvalue, g = kwclass)
  pvalue[i] <- dfB[3]
}

#Adjust p-value by Benjamini-Hochberg correction for multiple testing
pvalue_new <- round(p.adjust(pvalue, "BH"),3)

#Choose only features with adj. p-value < 0.05
indFeatures <- which(pvalue_new < 0.05 , pvalue_new)
sigfeat <- feature_list[indFeatures]

```
#Plot Univariate Plots

```{r Univariate Plots}

#Only use these plots with significant differences for features in sigfeat
#(have a significant Kruskal-Wallis score). Otherwise, comment out 
#stat_compare_means lines 312-317 to get basic plots.

i = 0
for (feature in feature_list){
  i = i+1
  testtest <- dfBox[dfBox$features == feature,] #pull only feature of interest
  
  p <- ggplot(testtest, aes(x = group, y = value, fill = group), color = group) + 
      scale_fill_manual(values = c('#8C02CC','#5192C3','#45C902','#045CFF')) +
      stat_boxplot(position = position_dodge(width = 0.75), 
                   geom = 'errorbar', width = 0.2) +
    # don't show outliers
     geom_boxplot(outlier.shape = NA) + 
    # add points for individual measurements
     geom_point(position = position_jitterdodge(), shape = 21, size = 1, 
                aes(fill = group), color = "black") + 
     ylab("z-scored value") + 
     xlab(feature) +
     stat_compare_means(method = "wilcox.test", #Mann-Whitney U
                        paired = FALSE,
                        #Which comparisons do you want to make?
                        comparisons = my_comparisons, 
                        #What star number do you want for each sig. level?
                        symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 
                                                         0.01, 0.05, 1), 
                                           symbols = c("****", "***", "**", 
                                                       "*", "ns")), 
                        hide.ns = TRUE) +
     theme_classic() +
     theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

    #print to pdf- change to be appropriate to your own directory
    pdf(paste0("../MalariaVax/Figures/Univariate_", feature, ".pdf"), 
           width = 8, height = 5) 
    print(p)
    dev.off()
}
```

#Multivariate plotting
PCA and PLS-DA (using the package ropls) and visualization.

To compare Group 3 vs Group1/2/4 combined, run this box
```{r G3 v G124}
y_124 <- as.character(y)

#Rename all subjects in groups 1,2,4 as members of combined group, and members
#of group 3 as members of that group
y_124[which(grepl("1|2|4", y_124))] <- "G124"
y_124[which(grepl("3", y_124))] <- "G3"
#subgroup-specific X and y values
y_subgroup <- as.factor(y_124)
X_subgroup <- Ximp
```

To compare Group 3/4 combined vs Group 1/2 combined, run this box
```{r G34 v G12}
y_12 <- as.character(y)
#Rename all subjects in groups 1, 2 as members of combined group, and members
#of groups 3,4  as members of combined group
y_12[which(grepl("1|2", y_12))] <- "G12"
y_12[which(grepl("3|4", y_12))] <- "G34"
#subgroup-specific X and y values
y_subgroup <- as.factor(y_12)
X_subgroup <- Ximp
```

To compare Group 3 vs Group 4, run this box
```{r G3 v G4}
y_34 <- as.character(y)
#Rename all subjects in group 3 as members of that group, and members
#of group 4  as members of that group. Drop all others.
y_34[which(grepl("3", y_34))] <- "G3"
y_34[which(grepl("4", y_34))] <- "G4"
Ximp <- Ximp[which(grepl("3|4", y_34)),]
y_subgroup <- as.factor(y_34[which(grepl("G3|G4", y_34))])

#subgroup-specific X and y values
y_subgroup <- droplevels(y_subgroup)
X_subgroup <- as.matrix(Ximp)
```

#PLS-DA, all features
Visualize group separation with PLS-DA, using all features
```{r PLSDA vis}

#Define colors based on feature class and group
my_colors <- list(group = c('G3' = '#45C902','G124' = '#2D3880', 
                            'G12' = '#8C02CC', 'G34' = '#FFA704', 
                            'G4' = '#045CFF'), 
                  feature_class = c('glycan' = '#AB6FC9', 'function' = '#5192C3',
                                    'titer' = '#6FB06D', 'FcR' = 'red'))

# general options for plotting
opts_plot <- list(df_features = df_features,
                  loading_alpha = 1, # transparency for the loadings
                  score_alpha = 1, # transparency for the scores
                  LV_ind = c(1,2), # which LVs to plot
                  n_LV = 2, # number of latent variables
                  # according to which property the features should be colored
                  color_features = "group", 
                  colors = my_colors, #which colors do you want to use?
                  y_name = "group") #what should your subjects be named by

opts_plot2 <- list(df_features = df_features,
                  loading_alpha = 1, 
                  score_alpha = 1, 
                  LV_ind = c(1,2), 
                  n_LV = 2,
                  color_features = "feature_class", 
                  colors = my_colors,
                  y_name = "group") 

# Perform a simple PCA
model_pca <- pca_ropls(X_subgroup)
ropls::getSummaryDF(model_pca) #Get PCA summary diagnostics
plt_scores_pca <- visualize_ropls_scores(model_pca, y_subgroup, 
                                         options = opts_plot)
print(plt_scores_pca) #plot PCA scores

# Perform a PLS-DA and plot the scores and loadings
model <- train_ropls(X_subgroup, y_subgroup, options = opts_plot2)
ropls::getSummaryDF(model) #Get PLS-DA summary diagnostics
plt_scores <- visualize_ropls_scores(model, y_subgroup, options = opts_plot2)
print(plt_scores) #plot PLS-DA scores

plt_loadings <- visualize_ropls_loadings(model, options = opts_plot2)
print(plt_loadings) #plot PLS-DA loadings

#Set additional options required to color code enrichment 
#in the bar plot of the loadings

#Latent Variable 1
opts_plot$X <- X_subgroup
opts_plot$y <- y_subgroup
opts_plot$LV_ind <- 1 #which latent variable should be plotted?
opts_plot$mark_enrichment <- TRUE #will mark which group feature is enriched in
plt_loadings_bar <- visualize_ropls_loadings_bar(model, options = opts_plot)
print(plt_loadings_bar)

#Latent Variable 2
opts_plot2$X <- X_subgroup
opts_plot2$y <- y_subgroup
opts_plot2$LV_ind <- 2
opts_plot2$mark_enrichment <- TRUE
plt_loadings_bar <- visualize_ropls_loadings_bar(model, options = opts_plot2)
print(plt_loadings_bar)

```

#Cross Validation, all features
Validate full model, 5 folds and 10 trials. Also permute 5 fold, 100 trials
This process will take time to run, so allow for adequate time

```{r Validation of full model}
method = list(train = train_ropls,
              predict = predict_ropls,
              score = score_accuracy)

#How many total trials, permutation (pt) trials, and random feature (rf) trials?
opts = list(n_trials = 10, n_folds = 5, pt_trials = 100, rf_trials = 0)

return_vals1 <- validate_repeat(X_subgroup, y_subgroup, 
                                method, opts, n_trials = 10)
systemsseRology::visualize_validate(return_vals1)
u1 <- rowMeans((sapply(return_vals1, unlist)))[1]
#Average cross-validation score
u1
```

#Feature Selection Validation and Permutation testing

Visualize group separation with PLS-DA, using LASSO-selected features

```{r Select Features}

#Select features using LASSO feature selection, 100-rounds, keeping only
#features selected in 80+% of rounds
opts_sel <- list(n_trials = 100, threshold = 0.8, return_count = TRUE)
sel_features <- select_repeat(X_subgroup, y_subgroup, 
                              selector = select_lasso, 
                              options = opts_sel)
X_sel <- X_subgroup[, sel_features$sel_features] #retain only selected features

#To test the importance of FcRN on overall results, remove FcRN as a feature
#Uncomment this to explore importance
#X_sel <- X_sel[,which(colnames(X_sel) != "FcRN")]

# Perform a PLS-DA using the selected features and plot the scores and loadings
# Check number of latent variables and increase to 2 if <2 
#(for visualization purposes)
opts_plot2 <- list(df_features = df_features,
                  loading_alpha = 1, 
                  score_alpha = 1, 
                  LV_ind = c(1,2), 
                  n_LV = 2,
                  color_features = "feature_class", 
                  colors = my_colors,
                  y_name = "group") 
opts_model <- list(n_LV = 2,
                  color_features = "group", 
                  colors = my_colors,
                  y_name = "group",
                  y = y_subgroup)
#Perform PCA
model_pca <- pca_ropls(X_sel)
ropls::getSummaryDF(model_pca) #summary statistics

#Perform PLS-DA with only selected features
model <- train_ropls(X_sel, y_subgroup, options = opts_model)
ropls::getSummaryDF(model) #summary statistics

#Visualize PLS-DA scores with only selected features
plt_scores <- visualize_ropls_scores(model, y_subgroup, options = opts_model)
print(plt_scores)

#Visualize PLS-DA loadings with only selected features
plt_loadings <- visualize_ropls_loadings(model, options = opts_plot2)
print(plt_loadings)

#print to pdf, changing folder path to suite your environment
#Additionally, consider changing file name to correct group comparison
pdf("../MalariaVax/Figures/PLSDA_scores_G34vG12_selected.pdf", 
       width = 8, height = 5) 
print(plt_scores)
dev.off()
pdf("../MalariaVax/Figures/PLSDA_loadings_G34vG12_selected.pdf", 
       width = 8, height = 5) 
print(plt_loadings)
dev.off()

# set additional options required to color code enrichment in the bar plot of the loadings
#Latent Variable 1
opts_plot$X <- X_sel
opts_plot$y <- y_subgroup
opts_plot$LV_ind <- 1 #which latent variable should be plotted?
opts_plot$mark_enrichment <- TRUE #will mark which group feature is enriched in
plt_loadings_bar <- visualize_ropls_loadings_bar(model, options = opts_plot2)
print(plt_loadings_bar)
#Latent Variable 2
opts_plot2$LV_ind <- 2
plt_loadings_bar <- visualize_ropls_loadings_bar(model, options = opts_plot2)
print(plt_loadings_bar)

#plot of times selected vs feature, colored by feature name/class
df_features_sel <- df_features
row.names(df_features_sel) <- df_features$name
df_features_sel$feature_count <- sel_features$feature_count
plt <- ggplot(df_features_sel, aes(x = reorder(name,-feature_count), 
                                  y = feature_count, fill = feature_class)) + 
     geom_bar(stat = "identity", show.legend = TRUE) +
     geom_hline(yintercept = 80, color = "black") +
     scale_fill_manual(values = my_colors$feature_class) +
     ylab("Selection Frequency") +
     xlab("Feature") +
     theme_classic() +
     theme(axis.text.x = element_text(angle = -90, vjust = .8, hjust=0))

plt

```

#Validation, Permutation Testing on Selected Features

```{r Validation, Permutation of Selected Features}
#Set options for feature selection approach, using LASSO feature selection
opts_sel <- list(n_trials = 100, threshold = 0.8, return_count = FALSE)
select <- function(X, y) { return(select_repeat(X, y, 
                                                selector = select_lasso, 
                                                options = opts_sel)) }
method = list(select = select,
              train = train_ropls,
              predict = predict_ropls,
              score = score_accuracy)

#How many total trials, permutation (pt) trials, and random feature (rf) trials?
opts = list(n_folds = 5, pt_trials = 1, rf_trials = 1)
return_vals2 <- validate_repeat(X_subgroup, y_subgroup, 
                                method, opts, n_trials = 100)
visualize_validate(return_vals2)
#Average cross-validation scores across trials
u2 <- rowMeans(sapply(return_vals2, unlist))[1]
u2
u3 <- rowMeans(sapply(return_vals2,unlist))
u3

```

#Correlation Mapping as heatmap

```{r correlation heatmap}
#Spearman correlations between all features
corrMat <- cor(X_subgroup, method = "spearman")
# heatmap for full correlation matrix
pheatmap(-corrMat, cluster_rows = FALSE, cluster_cols = FALSE,        
         color = colorRampPalette(brewer.pal(n = 10, name = "RdBu"))(100),
         breaks = seq(1, -1, length = 100),
         cellwidth = 7, cellheight = 6, fontsize = 6, border_color = "NA") 

# keep only selected features and highly correlated features
colMax <- function(X) apply(X, 2, max)
corrMat <- corrMat[sel_features$sel_features, ]
corrMat <- corrMat[, which(colMax(abs(corrMat)) < 1 & colMax(abs(corrMat)) > 0.7)]
# heatmap for reduced correlation matrix
pheatmap(-corrMat, cluster_rows = FALSE, cluster_cols = FALSE,        
         color = colorRampPalette(brewer.pal(n = 10, name = "RdBu"))(100),
         breaks = seq(1, -1, length = 100),
         cellwidth = 20, cellheight = 20, fontsize = 8, border_color = "NA")

```

## Co-correlation network of features correlated to selected features

```{r Co-correlation network, message=FALSE} 
#Spearman correlations between all features
tidyCors <- X_subgroup %>% 
  correlate(method = "spearman") %>% 
  stretch()

#Test correlation significance and store p value in new p variable
tidyCors$p <- 0
for (ind in 1:dim(tidyCors)[1]) {
  tmp <- cor.test(X[,tidyCors$x[ind]], X[,tidyCors$y[ind]], 
                  method = "spearman", exact = FALSE)
  tidyCors$p[ind] <- tmp$p.value
}
#Multiple hypothesis correction via Benjamini-Hochberg
tidyCors$p <- p.adjust(tidyCors$p, method = "BH", n = length(tidyCors$p))

#Which features do you want to plot correlations to?
plotFeatures <- sel_features$sel_features #selected features

#Pull out only significant (p<0.05) and highly correlated 
#(abs(spearman corr) > 0.7) features
graphCors <- tidyCors %>% 
  filter(p < 0.05 & abs(r) > 0.7 & 
           (is.element(x, plotFeatures) | is.element(y, plotFeatures))) %>% 
  graph_from_data_frame(directed = FALSE)

#Visualization options for network graphs
layout <- create_layout(graphCors, layout = 'igraph', algorithm = 'in_circle') 
            #in_circle, nicely, with_kk are all good options
nodeColours <- vector(mode = "list", length = length(V(graphCors)$name))
    # selected features colored gray
nodeColours[is.element(V(graphCors)$name, plotFeatures)] <- 'gray' 
    # other features colored white
nodeColours[!is.element(V(graphCors)$name, plotFeatures)] <- 'white' 

pltGraph <- ggraph(layout) +
  geom_edge_link(aes(color = r), edge_width = .5) +
  guides(edge_alpha = "none", edge_width = "none") +
  scale_edge_colour_gradientn(colors = rev(colorRampPalette(brewer.pal(n = 11, 
              name = "RdYlBu"))(100)))+ #choose color palette for visualization
  geom_node_point(shape = 21, color = "gray", 
                  fill = nodeColours, size = 9, stroke = 0.5) +
  geom_node_text(aes(label = name, fontface = "bold"), size = 2, 
                 point.padding = NA,  box.padding = 0, force = 0.1, repel = T) + 
  theme(aspect.ratio = 1) +
  theme_graph(background = "white", base_family = 'Helvetica') 

plt <- plot(pltGraph)

#print to pdf, changing folder path to suite your environment
#Additionally, consider changing file name to correct group comparison
pdf("../MalariaVax/Figures/CorrelationNetwork_G3vG4_circle.pdf", 
       width = 15, height = 8) 
print(plt)
dev.off()
```
